---
title: "exercise-09"
output: html_document
date: "2024-03-06"
---

## Step 1

This exercise is working with the published dataset from Street et al. (2017).
```{r}
library(tidyverse) 
library(dplyr)
library(skimr) #To use the skim() function

f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/Street_et_al_2017.csv"
(d <- read_csv(f,col_names = TRUE)) #As a tibble, d
```


To generate the five-number summary (median, minimum and maximum and 1st and 3rd quartile values), plus mean and standard deviation, for each quantitative variable, I am using the skim() function from {skimr}.
```{r}
skimd <- skim(d) |> 
  filter(skim_type == "numeric") |>
  select(skim_type, skim_variable, numeric.p50, numeric.p0, numeric.p100, numeric.p25, numeric.p75, numeric.mean, numeric.sd)

names(skimd)[names(skimd) == "numeric.p50"] <- "median"
names(skimd)[names(skimd) == "numeric.p0"] <- "minimum"
names(skimd)[names(skimd) == "numeric.p100"] <- "maximum"
names(skimd)[names(skimd) == "numeric.p25"] <- "Q1"
names(skimd)[names(skimd) == "numeric.p75"] <- "Q3"
names(skimd)[names(skimd) == "numeric.mean"] <- "mean"
names(skimd)[names(skimd) == "numeric.sd"] <- "sd"

(as.tibble(skimd))
```

## Step 2

Plotting **brain size (ECV)** as a function of **social group size (Group_size)**:
```{r}
(p1 <- ggplot(data = d, aes(x = Group_size, y = ECV)) +
   xlab("Group Size") + 
   ylab("Brain Size (ECV)")+
   geom_point(na.rm = TRUE))
```

Plotting **brain size (ECV)** as a function of **longevity (Longevity)**:
```{r}
(p2 <- ggplot(data = d, aes(x = Longevity, y = ECV)) +
   xlab("Longevity") + 
   ylab("Brain Size (ECV)") +
   geom_point(na.rm = TRUE))
```

Plotting **brain size (ECV)** as a function of **juvenile period length (Weaning)**:
```{r}
(p3 <- ggplot(data = d, aes(x = Weaning, y = ECV)) +
   xlab("Juvenile Period Length") + 
   ylab("Brain Size (ECV)")+
   geom_point(na.rm = TRUE))
```

Plotting **brain size (ECV)** as a function of **reproductive lifespan (Repro_lifespan)**:
```{r}
(p4 <- ggplot(data = d, aes(x = Repro_lifespan, y = ECV)) +
   xlab("Reproductive Lifespan") + 
   ylab("Brain Size (ECV)")+
   geom_point(na.rm = TRUE))
```

## Step 3

Deriving the ordinary least squares regression coefficients, β1 and β0, for ECV as a function of social group size:

Note: 
β1 = cov(x,y)/var(x))
β0 = mean(y) - β1*(mean(x))
```{r}
d2 <- d |> drop_na("ECV","Group_size")

(β1 <- cov(d2$Group_size,d2$ECV)/var(d2$Group_size))
(β0 <- mean(d2$ECV) - β1 * mean(d2$Group_size))
```

## Step 4

Now, I can check my β1 and β0 derived in **Step 3** using the lm() function:
```{r}
(lm(ECV ~ Group_size, data = d2))
```
The regression coefficients returned by lm() match those derived in **Step 3**.

## Step 5

### Repeating the above analysis for catarrhines only:

Repeating **Step 3** for catarrhines only:
```{r}
catarrhines <- d2 |> filter(Taxonomic_group == "Catarrhini")

(cβ1 <- cov(catarrhines$Group_size,catarrhines$ECV)/var(catarrhines$Group_size))
(cβ0 <- mean(catarrhines$ECV) - cβ1 * mean(catarrhines$Group_size))
```

Repeating **Step 4** for catarrhines only:
```{r}
(lm(ECV ~ Group_size, data = catarrhines))
```

### Repeating the above analysis for platyrrhines only:

Repeating **Step 3** for platyrrhines only:
```{r}
platyrrhines <- d2 |> filter(Taxonomic_group == "Platyrrhini")

(pβ1 <- cov(platyrrhines$Group_size,platyrrhines$ECV)/var(platyrrhines$Group_size))
(pβ0 <- mean(platyrrhines$ECV) - pβ1 * mean(platyrrhines$Group_size))
```

Repeating **Step 4** for platyrrhines only:
```{r}
(lm(ECV ~ Group_size, data = platyrrhines))
```

### Repeating the above analysis for strepsirhines only: 
Repeating **Step 3** for strepsirhines only:
```{r}
strepsirhines <- d2 |> filter(Taxonomic_group == "Strepsirhini")

(sβ1 <- cov(strepsirhines$Group_size,strepsirhines$ECV)/var(strepsirhines$Group_size))
(sβ0 <- mean(strepsirhines$ECV) - sβ1 * mean(strepsirhines$Group_size))
```

Repeating **Step 4** for strepsirhines only:
```{r}
(lm(ECV ~ Group_size, data = strepsirhines))
```
Looking at these values sides by side...
```{r}
group <- as.tibble(c("Catarrhines","Platyrrhines","Strepsirhines"))
names(group)[names(group) == "value"] <- "Taxonomic Group"
beta1 <- as.tibble(c(cβ1,pβ1,sβ1))
names(beta1)[names(beta1) == "value"] <- "β1"
beta0 <- as.tibble(c(cβ0,pβ0,sβ0))
names(beta0)[names(beta0) == "value"] <- "β0"

(compare <- bind_cols(group,beta1,beta0))
```

The regression coefficients appear to differ among groups. One way to test this would be to use a permutation test to evaluate whether or not there was a statistically significant difference between the values.

INCOMPLETE ## Step 6
Calculating the standard error for the slope coefficient: the 95% CI, and the p value associated with this coefficient for the regression of ECV on social group size from my original **Step 3** and **Step 4**:

```{r}
#Calculating the standard error for the slope coefficient:
m <- (lm(ECV ~ Group_size, data = d2))

SSE <- sum((m$residuals)^2)

SSX <- sum((m$model$Group_size-mean(m$model$Group_size))^2)

MSE <- SSE/(nrow(d2)-2)

(seβ1 <- sqrt(MSE/SSX))
```

```{r}
#Calculating the 95% CI:

```


Checking these calculations by extracting the same information from the results of the lm() function:
```{r}
library(broom)
(tidy(m))
(confint(m,"Group_size",level=0.95))
```

The seβ1 is 0.3508, which matches the seβ1 I calculated by hand.
The 95% CI is [1.769874,3.156269].
The p-value is 7.259e-11.

INCOMPLETE: ## Step 7

INCOMPLETE: ## Step 8
